---
title: "Introduction to R"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Using R as command prompt

```{r}
5+6
5e10
```

## Using R to solve algebraic equations


```{r}
# Initial values
x = 10
y = 20

# Equation
magnitude = sqrt(x^2 + y^2)
vx = x/magnitude
vy = y/magnitude

print(vx)
print(vy)

```

## Using R and vector functions

```{r}
x = c(1,3,8)
x[2]
length(x)
```

## Exercise 1
```{r}
# Option 1
log(((3+2)*5)+6)
```
```{r}
# Option 2
x = 3+2
y = x*5
z = y + 6
log(z)
```
```{r}
# Create a character vector with 5 elements, then print the fourth element
vector = c(6,7,8,9,1)
print( vector[4] )
```
```{r}
vector2 = c(6,7,8,9)
print( vector2[1] + vector2[3] )

```

## Using matrices

```{r}
x = c(1,3,8)
y = c(3,4,5)

cbind(x,y)
```
```{r}
rbind(x,y)
```

```{r}
xy = rbind(x,y)
dim(xy)
```

```{r}
row = 1
column = 3
xy[row, column]
```


```{r}
# Negative index value
print(y)
y[-2]
```

## Retrive submatrices
```{r}
print(xy)
```

```{r}
# Get all the rows from the column 3
xy[ , 3]
```
```{r}
# Get all the columns from the row 1
xy[ 1 , ]
```

## Matrices operations
```{r}
xy
```

```{r}
# Scalar
xy * 10
```
```{r}
# Elementwise
x
y
x*y
```

```{r}
xy %*% t(xy)
```

## Exercise 2
```{r}
# plot hist of a random normal variable
randomValues = rnorm(1000, mean = 10, sd = 3)
hist(randomValues)
```

## Exercises 3
Use control flow to print all the values from 1 to 100 that are divisible by 3

```{r}
# Option 1: Using while and addition

i = 3
while ( i <= 100 ){
  print(i)
  i = i + 3
}
 

```

```{r}
# Option 2: Using while and multiplication

i = 1
# floor is a function used to truncate a decimal value
upperBound = floor(100/3)
while ( i <= upperBound ){
  print(i*3)
  i = i + 1
}
```

```{r}
# Option 3: Using for and multiplication

# floor is a function used to truncate a decimal value
upperBound = floor(100/3)
for(i in 1:upperBound){
  print(i*3)
}
```


Use control flow to print all the numbers in the Fibonacci series till its 20th element

```{r}
past = 0
present = 1

for ( i in 1:20){
  print(present)
  future = present + past
  past = present
  present = future
}

```

Use control flow to print if a numeric variable is a prime number 

```{r}

# Option 1: Without using functions

prime = TRUE
number = 11

for( i in 2:(number-1) ){
  # This if is used to check if the variable "bumber" is perfectly divisible by the variable "i"
  if (number %% i == 0 ) {
    # If the number can be divided by i, then it means that the number is not prime
    prime = FALSE
    # We use break to end the loop
    break
  }
}

answer = paste("Your number is a prime number?", prime)
print(answer)


```



```{r}

# Option 2: Using a function

isPrime <- function(number){
  for( i in 2:(number-1)){
    if ( number %% i == 0){
      return(FALSE)
    }
  }
  return(TRUE)
}

isPrime(11)
isPrime(12)

```

## Exercise 4
```{r}
data(iris)
ncol(iris)
```

```{r}
nrow(iris)
```

```{r}
dim(iris)
```

```{r}
head(iris)
```

```{r}
colnames(iris)
```

```{r}
summary(iris)
```

## Exploratory data analysis
```{r}
hist(iris$Sepal.Length)
```

```{r}
plot(density(iris$Sepal.Length))
```

```{r}
boxplot(iris)
```

```{r}
boxplot(iris, outline = FALSE)
```


## Filtering data
```{r}
data("iris")
library(dplyr)
filteredDatabase = filter( iris, Sepal.Width > 4)
```


```{r}
variableSlice = slice( iris, 10:15) 
slice( iris, 10:15) 
```












## Example how to write equations using latex notation

$$
x = \sum_{k=1}^K k \\
x = 1\\
y = 1
$$











