---
title: "5_Integrals"
author: "Roberto Cardenas"
date: "16/8/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Getting the first n-terms of the Taylor series

```{r}

library(pracma)
taylor(f = sin, x0 =  0, n = 6)
```

Using the integral functions included in pracma

```{r}
function1 <- function(x) {
  exp(-(x^2))
}

integral(function1, 0 , 0.5)
```

Using Monte Carlo approach with an evaluation range of 0 to 1

```{r}
fx = expression(exp(-(x^2)))

x <- runif(100000)
X <- eval(fx)
mean(X)
```

Using Monte Carlo approach with an evaluation range different than 0 to 1

First method: Using the transformation from x to y
integral range [0, 0.5]

```{r}
gu <- expression(exp(-((0.5*u)^2))*0.5)

u <- runif(10000)
U <- eval(gu)
mean(U)
```

Second method: Using x, but being careful of the bounds. The bounds (b-a) from the derivative value
```{r}
u <- runif(10000, 0, 0.5)
gu <- expression( exp(-(u^2)) )
U <- eval(gu)
mean(U*0.5) # <- el 0.5 comes from [b-a]dy = dx
```

Let's compare it with the integral method from R
```{r}
integral(function1, 0, 1/2)
```


Multivariate integrals with pracma

```{r}
function2 <- function(x, y){
  exp((x+y)^2)  
}

integral2(function2, 0, 1, 0, 1)

```

## Extra

Compute $$\pi$$ by using known areas

```{r}
# Number of random points
L <- 1e4
r = 1

# Generation of random points
X <- runif(L,-1,1)
Y <- runif(L,-1,1)

# Formula to satisfy / constraint
total <- (X^2 + Y^2) <= r^2

# Let's see how many points satisfy the constraint
# The 4 comes from the range(-1, 1) for both variables (square area)
(4*sum(total))/length(total)
```

Let's plot it 
```{r}
color <- (as.numeric(total)+1)
plot(X,Y, pch = 19, col = color)
```

The same idea, but only using one quarter of the circule

```{r}
L <- 1e6

X <- runif(L,0,1)
Y <- runif(L,0,1)

total <- (X^2 + Y^2) <= 1
(4*sum(total))/length(total)
```

```{r}
color <- (as.numeric(total)+1)
plot(X,Y, pch = 19, col = color)
```

