---
title: "HM_2_Statistical_Tests"
author: "Jacobo Posada Hoyos A00829481 - Diego Fernando Cabrera Castellanos A00829475 - ITESM Campus Querétaro"
date: "6/9/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
### 1. Statistic Test
## Charging the data
```{r}
# Cleaning the console and Global Enviroment
rm(list = ls())
cat("\014") 


# Calling the useful libraries
library(dplyr)

# Importing the database
dataImp = read.csv("Dataset.csv")
dataImp = select(dataImp,2,4,6,8,9)

filteredData = filter(dataImp, Variable == 'Drug use disorders' 
                      & Measure == 'Deaths per 100 000 population (standardised rates)')

filtData2003 = filter(filteredData,Year == 2003)
Value2003 = filtData2003$Value
filtData2014 = filter(filteredData,Year == 2014)
Value2014 = filtData2014$Value
```

## Getting the pdf of the variables
```{r}
hist(Value2003)
hist(Value2014)
plot(density(Value2003))
plot(density(Value2014))
```

## Performing the Mann Whitney- U Test.
```{r}
#Both of vectors do not follow a normal probability distribution
wilcox.test(Value2003,Value2014, paired = F, alternative = "two.sided",conf.int = 0.95)
```

## Executing the Bootstraping Method
```{r}
len2003 = length(Value2003)
len2014 = length(Value2014)
numIt = 100
pValues1 = numeric(numIt)

for(i in 1:numIt){
  if(len2003>0 & len2014>0){
    set.seed(len2003)
    sample2003 = sample(Value2003,len2003)
    set.seed(len2014)
    sample2014 = sample(Value2014,len2014)
    pValues1[i] = wilcox.test(sample2003,sample2014, paired = F, 
                              alternative = "two.sided",conf.int = 0.95)$p.value
  
    len2003 = ceiling(0.75*len2003)
    len2014 = ceiling(0.75*len2014)
  }
}

hist(pValues1,main ="Bootstraping Method performing Mann-Whitney U Test", 
     xlab = paste("Mean P-Values: ",toString(mean(pValues1))),col=rgb(0.1,0.1,0.1,0.5))
box()
```

## Interpretation: 
The statistical test was made for 2 events which contents information about the percentage of death per 100000 population, in several countries, owing to drug use disorders in 2 years specifically: 2003 & 2014. 
The hyphotesis works around wether the difference of death is significantly different in both years. 
After analyzing the probability density function of 2 events, teh distribution does not fulfill a normal one, therefore it was convenient using a statistical test wich was non-parametric (like this test). So, the chosen was Mann Whitney U Test whitch 95% as picked confidence interval. The p-value provides the samples range that produce a considerably difference and, for this case, it is just around of 15 % (The obtained p-value is 0,15 approximately).
Thus, at a significance level of 0.05, the difference between the percentage of deaths (per each 100000 habitants) does not seem being significantly different in 2003 and 2014 by drug use disorders. 

### 2. Mann-Whitney U Test - Function
```{r}
mannWhitneyU <- function(d1, d2){
  
  ## Saving the samples in data frames
  data1 <- data.frame(param = d1, AB = "A" )
  data2 <- data.frame(param = d2, AB = "B" )
  ## Combining the data frames
  x = rbind(data1,data2)
  df <- data.frame("D" = x[,1], "AB" = x[,2] )
  
  # Sorting and Mixing the data in one variable
  dfSort <- df[do.call(order, df), ]
  
  # Calculating the rank of each data
  dfTotal = data.frame(dfSort, "Rank" = rank(dfSort[,1]))
  
  ## Now we do the statistical calculation 
  # U1,2 = n1*n2 + (n1,2(n1,2+1))/2 - R1,2
  # U = min(U1,U2) 
  
  
  ## Calculating R1
  RA = dfTotal[,2,] == "A"
  R1 = dfTotal[,3,]
  R1 = sum(as.numeric(as.character(R1[c(RA)])))
  
  ## Calculating R2
  # So, we can know the value R2
  ## Calculating n1 and n2
  n1 <- dim(data1)[1]
  n2 <- dim(data2)[1]
  N = n1+n2
  R2 = abs(((N*(N+1)/2) - R1))

  #Now we can obtain U1 and U2
  U1 <- (n1 * n2) + (n1*(n1 + 1) / 2) - R1
  U2 <- (n1 * n2) + (n2*(n2 + 1) / 2) - R2
  
  # Now, calculating U
  U  <- min(c(U1,U2))
  
  # Check for U values <25
  if(U<25){
    pValue = pwilcox(U,n1,n2)
  }
  else{
    z <- (U-(n1*n2)/2)/(sqrt((n1*n2*(n1+n2+1))/12))
    pValue = pnorm(z)
  }
  cat("               U Mann-Whitney Test\n\n ")
  cat("U = " ,U,",","p-value = ", pValue,"\n")
}

### Now, we can perform the test for the pima data.
# Libraries
library(faraway)

# Obtaining Pima dta
data(pima)

# Filtering 0-values
filteredPima = filter(pima, glucose > 0,diastolic > 0, triceps > 0, insulin > 0, bmi > 0)

# Selecting the parameter to make the test
filteredPimaD1 = select(filter(filteredPima, test == 1), diabetes)
filteredPimaD2 = select(filter(filteredPima, test == 0), diabetes)

# Our U Mann-Whitney Function
mannWhitneyU(filteredPimaD2,filteredPimaD1)

# R Mann-Whitney Function 
wilcox.test(filteredPimaD2$diabetes,filteredPimaD1$diabetes, paired = FALSE)

# R t Test Function
t.test(filteredPimaD2$diabetes, filteredPimaD1$diabetes)

```





