---
title: "Final_ej1"
author: "Equipo"
date: "4/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(1111)
```

```{r}
rm(list = ls())
library (vcd)
library (nortest)
dataset = read.csv("Weather.csv")
summary(dataset)

```
```{r}
library(dplyr)


dataset = filter (dataset, PRCP != c(""))


dataset$PRCP <- as.numeric(as.character(dataset$PRCP))

```

```{r}

summary(dataset)

```


```{r}


filtered_data = filter (dataset, PRCP != c( "NA"))

```

```{r}
summary(filtered_data)

```
#After filtering we proceed with ANCOVA because we count with countinuous variables and categorical variables

```{r}


test=lm(PRCP~MaxTemp+MinTemp+MO, filtered_data)

summary.lm(test)


```

```{r}
filtered_data_mar = dplyr::filter(filtered_data, MO == 3)
filtered_data_oct = dplyr::filter(filtered_data, MO == 10)

```
```{r}
plot(density(filtered_data_mar$PRCP), col = "blue")
lines (density(filtered_data_oct$PRCP), col = "red")
boxplot(filtered_data_mar$PRCP, filtered_data_oct$PRCP)

```



```{r}

ad.test(filtered_data_mar$PRCP)
ad.test(filtered_data_oct$PRCP)
ks.test(filtered_data_mar$PRCP,filtered_data_oct$PRCP)


```
#According to the graphics and the results of the normality test the data does not behave normally because of this The test to be performed will be Mann-Whitney also known as two-sample Wilcoxon

```{r}
pval=numeric(100)
var<-1:100

test<-0
for (i in var) {
  bs1<-sample(filtered_data_oct$PRCP,25,replace = TRUE,prob = NULL)
  bs2<-sample(filtered_data_mar$PRCP,20,replace = TRUE,prob = NULL)
  pval[i]=wilcox.test(bs1,bs2, paired = F, alternative = "two.sided",conf.int = 0.95)$p.value
  
  
  if(pval>=0.05){
    test=test+1}
}
  hist(pval)
  mean(pval)
  plot(density(pval))

```

# Conclusion

According to the data there is not enough evidence to reject the null hypothesis, which means that with a 95 % of confidence that there is no significant difference between drugs use disorders as cause of dead from 2003 and 2014. 

