---
title: "R Notebook"
output: html_notebook
---


Q1: Use the gene data in the below link
http://web.stanford.edu/~hastie/CASI_files/DATA/leukemia_big.csv
	Perform 1000 nonparametric bootstrap replications of (ALL) ̅=0.752. You can use program bcanon from the CRAN library “bootstrap” 
	Do the same for (AML) ̅=0.95.
	Plot histograms of the results, and suggest an inference.
Suppose that there were no differences between AML and ALL patients for any gene, so that 

t=((ALL) ̅-(AML) ̅)/(sd) ̂ 

exactly followed a student-t distribution with 70 degrees of freedom in all 7128 cases. About how big might you expect the largest observed t value to be? Hint: 1/7128 = 0.00014. 



```{r}
library(ggplot2)
library(ggpubr)
theme_set(theme_pubr())
library(dplyr)
library(tidyverse)
all<-read.csv("C:/Users/Alejandro/Documents/1 ITESM/MCI/2do semestre/Nueva carpeta/Final/ALL.csv",  header = TRUE, stringsAsFactors = FALSE, na.strings = "NA")
```
```{r}
all<-select(all, -c(Columna1))

g136 <- select(all, X136)
```


```{r}

library(bootstrap)
g136<-as.numeric(unlist(g136))
theta<- function(g136){mean(g136)}

g136_boot <- bcanon(g136,1000, theta = theta)

g136_boot$u

hist(g136_boot$u)

```


```{r}
aml<-read.csv("C:/Users/Alejandro/Documents/1 ITESM/MCI/2do semestre/Nueva carpeta/Final/AML.csv",  header = TRUE, stringsAsFactors = FALSE, na.strings = "NA")
```

```{r}
aml<-select(aml, -c(Columna1))

g136_aml <- select(aml, X136)
```

```{r}

library(bootstrap)
g136_aml<-as.numeric(unlist(g136_aml))
theta<- function(g136_aml){mean(g136_aml)}

g136_boot_aml <- bcanon(g136_aml,1000, theta = theta)

g136_boot_aml$u

hist(g136_boot_aml$u)
```
#3

```{r}
library(stats)
library(bootstrap)


# aml_boot_mat<-matrix(, nrow=47, ncol=7128)
# aml_boot_vec<-vector()
# 
#   aml_boot <- bcanon(aml[,1],1000, theta = theta)
#   aml_boot_mat[[,1]] <-aml_boot$u
# aml_boot_vec[1]<-aml_boot$u
# aml_boot_vec<-vector()
# for (j in 1:7128){
#   aml_boot <- bcanon(aml[,j],1000, theta = theta)
#   aml_boot_vec[j]<-aml_boot$u
#   
# }
# aml_t<-g136_boot_aml$u
# all_t<-g136_boot$u

t_vec<-vector()
for (j in 1:7128){
  t<-t.test(aml[,j], all[,j])
  t_vec[j]<-t$statistic
}


plot(t_vec)
t_vec_boot <- bcanon(t_vec,1000, theta = theta)
qqnorm(t_vec_boot$u)
```


#6



```{r}
med<-read.csv("C:/Users/Alejandro/Documents/1 ITESM/MCI/2do semestre/Nueva carpeta/Final/med.csv",  header = TRUE, stringsAsFactors = FALSE, na.strings = "NA")
med
obs<-med

```
```{r}
#Parametrize data to obtain vales between 0 and 1
observed = med$y
maximo = max(obs$y)
med$y<-med$y/maximo

str(med)
sum(is.na(med$y))
logita<-glm(y~.,data=med, family = "binomial")
logita
summary(logita)

prediction = predict(object = logita, type = "response")

response = prediction*maximo
res <- trunc(response)

res
```




#7
(a) Fit the Poisson regression model to below data: 
https://web.stanford.edu/~hastie/CASI_files/DATA/galaxy.txt
(b) Plot the Poisson deviance residuals. 
(c) Where does the fit seem poor?
(d) How might you add to model to get a better fit? 



```{r}
ref<-read.csv("https://web.stanford.edu/~hastie/CASI_files/DATA/galaxy.txt",  header = FALSE, stringsAsFactors = FALSE, na.strings = "NA")
gx<-read.csv("C:/Users/Alejandro/Documents/1 ITESM/MCI/2do semestre/Nueva carpeta/Final/gx.csv",  header = TRUE, stringsAsFactors = FALSE, na.strings = "NA")
gx

```
```{r}
#Parametrize data to obtain vales between 0 and 1
observed = gx$gx
# maximo = max(data_tool$CP)
# data_tool$CP<-data_tool$CP/maximo

str(gx)
sum(is.na(gx$gx))
logita<-glm(gx~.,data=gx, family = "poisson")
logita
summary(logita)


#Model with only the 20 variables that affect the most


c<-sort(logita$coefficients)
c

prediction = predict(object = logita, type = "response")

prediction = trunc(prediction)

pred_mat <- matrix(, nrow = 18, ncol = 15)
rownames(pred_mat)<- c("21.38", "21.13", "20.88", "20.63", "20.38", "20.13", "19.88", "19.63", "19.38", "19.13", "18.88", "18.63", "18.38", "18.13", "17.88", "17.63", "17.38", "17.13")

colnames(pred_mat)<- c("-1.56", "-1.44", "-1.31", "-1.19", "-1.06", "-0.94", "-0.81", "-0.69", "-0.56", "-0.44", "-0.31", "-0.19", "-0.06", "0.06", "0.19")

c=1

for (j in 1:15) {
  for (i in 1:18){
      pred_mat[i,j]<-prediction[c]
      c <- c+1
  }
  
}

pred_mat
```
```{r}
estimates <- fitted(logita)
res.dev1 <- residuals(logita, type = "deviance")
res.dev2 <- residuals(logita)
poisson.dev <- function (y, mu) 
    # unit deviance
    2 * (y * log(ifelse(y == 0, 1, y/mu)) - (y - mu))
res.dev3 <- sqrt(poisson.dev(prediction, estimates)) * 
        ifelse(prediction > estimates, 1, -1)
all.equal(res.dev1, res.dev2, res.dev3)

qqnorm(residuals(logita))
plot(residuals(logita))

print(sd(resid(logita)))
```

